Molecular dynamics simulation of synchronization in driven particles

### 2021-01-07 manuscript/manuscript_ajp_2020.tex
fixed language in intro/abstract and rearranged for clarity

\begin{abstract}
  %What problem did you study and why is it important?
  Particles
  %confined to a narrow channel 
  driven
  across a washboard potential energy landscape
  exhibit a variety of synchronization effects.
  %in
  %  Numerical model
  We discuss a simple
  numerical model that
  %exhibits
  provides insight
  to synchronization behavior
  %and
  %atoms on surfaces.
  %voltage driven superconducting vortices
  %confined in a Josephson junction,
  %and ac and dc driven
  %charge and spin density waves.
  %We present the basics
  using
  molecular dynamics simulations
  for particles moving through
  a viscous liquid.
  This model reproduces behaviors observed
  in experimental systems 
  of interest to condensed matter physics 
  including 
  magnetically driven colloidal particles
  confined by light-fields.
  %
  %applied to single and multiple particle systems.
  %so we model the dynamics of each particle
  %with 
  %overdamped equations of motion.
  We demonstrate how to 
  visualize
  the system with
  numerical representations of the landscape
  and
  animations of the particle motion.
  Our results show a variety of synchronization effects 
  in single and multi-particle systems, 
  %and detect synchronization using
  %measurements of
  which we characterize with
  plots of 
  particle velocity as a function of applied driving force,
  and phase diagrams of position versus velocity.
  %We include sample code
  %and exercises for students
  %with 
  %opportunities
  %to reproduce our results and propose
  %new numerical experiments.
  %With only a few particles in two-dimensions,
  %the simulation runs quickly,
  %making this an appropriate model for undergraduates to explore.  
\end{abstract}
% AJP requires an abstract for all regular article submissions.
% Abstracts are optional for submissions to the "Notes and Discussions" section.

@@

\section{Introduction} %
%
%Outline
%------------------------------------------------------------------
%synchronization = widest audience possible
Synchronization is a universal phenomena
in which individual oscillators adjust frequency due
to external stimulus \cite{Pikovsky2003}.
In phase oscillations
%Synchronous behaviors where oscillators move in-phase
%(i.e. in time motions)
are observed
in many everyday systems
such as the
flickering candle flames coupled by temperature fluctuations \cite{Okamoto2016},
vibrations of singing wineglasses \cite{Arane2009}, 
and metronomes coupled through a supporting platform \cite{Jia2015}.
Biological systems benefit from cooperative
synchronization such as 
birds coordinating wing flaps
to optimize energy use during flight \cite{Portugal2014},
frogs alternating croaking patterns \cite{Aihara2014},
and humans clapping in time with music \cite{Tranchant2016}.
At a cellular level, 
neurons simultaneously fire in cardiac muscle \cite{MartinHall1999}
and brain tissue \cite{Singer1999}.

%Christiaan Huygens 
%[has been adapted into many useful technologies...]
A particular form of 
synchronization is phase-locking or mode-locking,
which first appeared in the scientific literature
with 
Huygens' 1665 experiments on
the 
motions of two wall-mounted pendulum clocks.
Huygens demonstrated the tendency of
the pendula 
%two periodic oscillators in close proximity
to swing in time %, i.e. to synchronize,
due to interactions through the wall,  
no matter the original phase of the clocks \cite{Bennett2002}.
%phase locking
Dynamical systems %with multiple frequencies
exhibit phase-locking 
when oscillators with different rhythms couple
so the frequency ratio or mode 
is an integer number \cite{Bak1986}.
%Dynamical mode locking
%The phenomena
Synchronization 
is often studied with simple computational models
and can be visualized 
with phase plots or Lissajous figures.
In phase space 
a mode-locked system is confined to 
a closed loop in a 
parametric space showing the relationship between two periodic functions
with patterns determined by the mode, % frequency ratio,
first reported in 1857 \cite{Lissajous1857}
and easily generated with a computer code 
or oscilloscope \cite{Tong1997}. 
%with a frequency generator,
%where two combined frequencies at perpendicular orientations
({\it As we will demonstrate in phase plots in Figs. X, Y, Z})

External forcing can cause or regulate 
synchronization. % is often achieved via external forcing
%such as when
%single particle oscillators
For instance a
%pushing a child on a swing at the high point on her path, or
pacemaker (an electrical generator)
pulses to regulate a heart beat
or controlling the pattern of flashing
fireflies with a pulsed light \cite{Agrawal2013}.
These behaviors can be modeled with
numerical studies %can be performed
where
a driving force is applied to simple harmonic system,
either
coupled oscillators %\cite{Kuramoto193?}
or a single particle driven across a periodic potential landscape.
A single particle particle oscillator
within a potential well is 
much like a skateboarder in a half-pipe or
a child on a swing
moving back and forth. % through equilibrium.  % \cite{}.
Single particle studies 
are useful to understand synchronization
in the absence of collective effects.
With a constant or 
dc drive the landscape modulates 
the particle velocity, below some threshold below
the dc force is not strong enough to push the particle
across a potential maximum so the average velocity is zero,
a phenomena refered to as pinning \cite{Reichhardt2017},
and dissipating energy from the system at all driving forces.
A sinusoidal or ac drive creates mode-locking,
where the average particle velocity
is fixed for a range of dc drive forces \cite{Reichhardt2015}.

@@

Complex dynamics can be studied with
colloid particles with experiments and simulation.
In experiment, colloids are typically spheres of 
plastic suspended in
highly de-ionized water or silica in suspended in organic solvent.
%experimental - colloids specifically
The spheres are a few microns in size
and move at rates of microns per second,
so a system can be imaged
with a traditional optical camera
due to the large size and slow speed of the particles.
Individual colloids exhibit 
Brownian motion,
where particles change direction in a seemly random fashion
due to unseen 
collisions 
with particles comprising the suspending fluid \cite{}.
Brownian motion is a temperature and viscosity dependent
phenomena. \cite{}
[elaborate - move to associated problems?].

@@

%Given the prevalence of synchronization,
%it is the subject of many computational models.
Here we perform numerical studies 
on the synchronized dynamics
of confined particles driven over
a washboard shaped potential energy landscape.
We chose this model for its
relevance to condensed matter systems
and ease of simulation.
%We examine
The period of the substrate
can be used to control the
intrinsic velocity of the dc driven particle,
and the applied ac drive can cause mode-locking
which appears as
Shapiro steps in the force-velocity relationship.

%---------------------------------------------------------------
%kinks and nanofriction

%-----------------------------------------------------------------
%transition from synchronized to chaotic
%describe types of physical systems - particle types
%%In this work,
%we perform 
%numerical simulations of confined, driven particles
%to model a variety of physical phenomena.
%-----------------------------------------------------------------------
%paper outline
%In the following paper
We describe
our molecular dynamics model for a single particle in Section~\ref{sec:MD}.
%In this section, we restrict our discussion to a single particle,
%and include additional information
%for multiple particles in Section~\ref{sec:sync}.
%including various  and confining environments.
We summarize
our results 
%in Section~\ref{sec:results} 
including synchronized motion of a single confined particle
driven across a periodic landscape in 
Section~\ref{sec:one}.
In Section~\ref{sec:sync}
we 
extend the model to include multiple particles,
and show how confined 
interacting particles respond to the same applied driving force.
%including stationary propagation of high density kinks
%in Sec.~\ref{sec:kink}.
%We include code to simulate
%and visualize the landscape and dynamics in this section
%and supplementary material.
%we
%demonstrate how uniform environments and applied forces
%and create synchronized flow patterns.
%We present these results using standard tools of non-linear oscillators
%such phase diagrams of velocity versus position.
%PHASE DIAGRAMS SHOWING COUPLING, OR QUASI-PERIODIC COUPLING
%INDICATE COOPERATIVE MOTION IN THE Q1D SUBSTRATE
%In Section~\ref{sec:quasiperiod},
%we show how an aperiodic landscape modifies the particle dynamics.
%Finally we explore the transition to chaotic dynamics in 
%Section~\ref{sec:chaos}
%and conclude 
%in Section~\ref{sec:conclusion}. 
We conclude with
exercises for interested students 
in Section~\ref{sec:problems}.

@@

In Fig.~\ref{fig:0} we
show the relationship between the applied force and particle
position % as a function of time.
for 
%We drive
a single particle driven across a periodic landscape. 
%along the y-direction
%({\it redundant, just refer back to eq. in Model?})
%and $V_{0y}=0.2$.
%This is illustrated in Fig.~\ref{fig:1}(a) where
%the red (blue) regions show local maxima (minima).
%
The numerical implementation of the landscape 
is calculated with Eq.~\ref{eq:ysubstrate} as 
\begin{equation}
  \label{eq:force}
  F^l_y(y) = -A_{p} \sin{(N_p \pi y / L)} 
\end{equation}
where the force is scaled with parameter $A_{p} = 0.1$ force units 
%in simulation units for force. % $E_0/a_0$.
%The substrate is described in 
%\begin{equation}
%  \label{eq:landscape_potential}
%  U_l(y) = U_{0y} \cos{(N_p \pi y / L)} 
%\end{equation}
with $N_p=20$ %troughs in the landscape
and spatial period $\lambda = 1.825 a_0$.
%
In Fig.~\ref{fig:0}(a)
we plot the applied driving force defined in Eq.~\ref{eq:drive}
as a function of time.
The time independent constants are 
$F^{AC}=0.2$ and $f=0.01$ cycles per time unit.
The temporal period of the driving force is
$T = 1/f = 100$.
To avoid transient oscillations at early times, 
the dc driving force is initially zero   
and 
slowly increased %$F^{DC}$.
at a rate of $\Delta F^{DC} = 0.01$ every $\Delta t = 4000$ integer timesteps 
to the constant value $F^{DC}=0.1$.
In Fig.~\ref{fig:0}(a) we plot $F^{DC}$
in blue to show the increase in 
$F^{DC}$ for the first $40$ time units.
In orange we show 
the combined ac and dc drive $F^D(t)$,
where 
the magnitude of the ac applied force is initially zero.
%since $\sin{(0)}=0$.
The vertical dashed lines in panels (a) and (b) coincide
with the condition $\sin{(\omega t)} = 1$,
where the applied force is maximum once $F^{DC}$ has become constant.

@@

The slope of $y/\lambda$
is proportional to the net force on the particle (See Eq.~\ref{eq:motion}).
This can be seen in 
Fig.~\ref{fig:0} %we demonstrate this relationship,
where local extrema appear in $y/\lambda$
when $F_{D}(t) \approx 0 $. 
When $F^D(t) > A_p$, a particle can 
overcome the barrier height of the landscape maxima,
and 
the particle hops between minima in the energy landscape.
In this example
we note 
the particle moves
in the positive y-direction
through $\Delta y = 2 \lambda$ in one time period $\Delta t = T$.
When $F_{D}(t) < 0 $
the particle moves in the negative y-direction 
for a short time.
%reaching a minimum position
%of $y/\lambda = 1$
%over the total period $T = 100$.
The average velocity $\bar{v}_y$
is the average displacement $\Delta y = y(t_0+T) - y(t_0)$
over the period of the driving force.
%In Fig. ~\ref{fig:0}
%the average displacement during one time period is a single wavelength
%of the substrate $y(t_0+T) - y(t_0) = \lambda$.
Thus the average velocity 
is $\bar{v}_y = 2 \lambda f$ where $T = 1/f$.

[add example where particle moves backward].

%single DC case with figure
%dc_current 0.01
%ac_current 0.2
%ac_frequency 0.0001
%fp_y 0.1
\begin{center}
\begin{figure}[h!]
\centering
\includegraphics[width=\columnwidth]{single_particle.pdf}
\caption{(a) The applied driving force $F^D(t)$ (orange) in Eq.~\ref{eq:drive}
  where the parameter $F^{DC}$ (blue) is slowly increased from zero to a constant value
  as described in the text.  (b) 
  The $y-$position of the driven particle as a function of time %of a single driven particle
  normalized by the period of the substrate $\lambda$.
  }
\label{fig:0}
\end{figure}
\end{center}

@@

We sample the average velocity $\langle v_y \rangle $
as a function of $F^{DC}$.

The hopping pattern of the driven particle is periodic,
and could be achieved over a range of $F^{DC}$.  %, as we will demonstrate in Fig.~\ref{fig:1}
We explore the ranges of periodic hopping patterns
by increasing $F^{DC}$ as a function of time,
as shown in 
%
Fig.~\ref{fig:1}.
%we show the result increasing the 
%force to the particle as a function of time.  
%increased DC
%We apply an external applied force, 
%with 
%a constant $F^{AC}$ with frequency $\omega = 2\pi f$
%then slowly increase $F^{DC}$
%at a rate of $\Delta F^{DC} = 0.001$ every $\Delta t = 4000$ integer timesteps
%or $40$ time units.
%this seems fast, but it is slow enough to estimate
%the velocity accurately.
%this is not be so for a multiple particle system!
The particle 
achieves a variety of oscillation modes as a function of $F^{DC}$.
A mode is a periodic pattern of hops
with a constant average particle velocity, $\bar{v}_{y}$
over a range of driving forces $F^{DC}$.
We illustrate mode-locking in 
the velocity-force plot in Fig.~\ref{fig:1}(b).
Here $\bar{v}_{y}$ is increasing in non-uniform steps,
with a quantized height of
$\bar{v}_{y} = n \lambda f$,
where $n$ is an integer.  
%$\lambda = L/N_p = 36.5/20 = 1.825$ is the spatial period, or wavelength
%of the landscape,
%and $f = 0.01$ cycles per time unit.

@@

We illustrate the hopping pattern in Fig.~\ref{fig:1}(b)
and 
show the dynamics 
in supplementary materials \cite{supp1}.

\begin{center}
\begin{figure}[h!]
\centering
\includegraphics[width=\columnwidth]{single}
\caption{\textbf{(a)} The particle is driven with a constant amplitude $F^{AC}$ and frequency $\omega$ through a periodic spatial potential landscape.  The landscape is represented with a colormap where blue are minima and red are maxima in the potential. \textbf{(b)} An average particle velocity in the y-direction $\bar{v}_{y}$ as a function of a constant driving force $F^{DC}$. In the animation available in Ref.~\cite{supp1} the magenta dot represents the average velocity of the particle $\bar{v}_{y}$ at which the particle in Fig. 1(a) is moving.}
\label{fig:1}
\end{figure}
\end{center}

@@

    \bibitem{Agrawal2013} D. K. Agrawal, J. Woodhouse, and A. A. Seshia, Physical Review Letters {\bf 111}, 084101 (2013).

>>>>

\begin{abstract}
  %What problem did you study and why is it important?
  Particles
  %confined to a narrow channel 
  driven
  across a washboard potential energy landscape
  move in 
  synchronized patterns.   
  %exhibit a variety of synchronization effects.
  %in
  %  Numerical model
  We discuss a simple
  numerical model that
  %exhibits
  provides insight
  to synchronization behavior
  %and
  %atoms on surfaces.
  %voltage driven superconducting vortices
  %confined in a Josephson junction,
  %and ac and dc driven
  %charge and spin density waves.
  %We present the basics
  using
  molecular dynamics simulations
  for particles moving through
  a viscous liquid.
  This model reproduces behaviors observed
  in experimental systems 
  of %interest to condensed matter physics %including 
  magnetically driven colloidal particles
  confined by light-fields.
  %
  %applied to single and multiple particle systems.
  %so we model the dynamics of each particle
  %with 
  %overdamped equations of motion.
  Our results show a variety of synchronization effects 
  in single and multi-particle systems, 
  %and detect synchronization using
  %measurements of
  which we characterize with
  plots of 
  particle velocity as a function of applied driving force,
  and phase diagrams of position versus velocity.
  %We demonstrate how to 
  %visualize
  %the system with
  %numerical representations of the landscape
  %and
  %animations of the particle motion.
  %We include sample code
  %and exercises for students
  %with 
  %opportunities
  %to reproduce our results and propose
  %new numerical experiments.
  %With only a few particles in two-dimensions,
  %the simulation runs quickly,
  %making this an appropriate model for undergraduates to explore.  
\end{abstract}
% AJP requires an abstract for all regular article submissions.
% Abstracts are optional for submissions to the "Notes and Discussions" section.

@@

\section{Introduction} %
%
%Outline
%------------------------------------------------------------------
%synchronization = widest audience possible
Synchronization is a universal phenomena
in which individual oscillators adjust frequency due
to external stimulus \cite{Pikovsky2003}.
Many everyday systems
exhibit in-phase oscillations of components 
such as the
flickering patterns of
candle flames coupled by temperature fluctuations \cite{Okamoto2016},
vibrations of singing wineglasses \cite{Arane2009}, 
and metronomes coupled through a supporting platform \cite{Jia2015}.
Biological systems benefit from cooperative
synchronization -- %such as 
birds coordinate wing flaps
to optimize energy use during flight \cite{Portugal2014},
frogs alternate croaking patterns \cite{Aihara2014},
humans clap in time with music \cite{Tranchant2016}
and 
at a cellular level, 
neurons simultaneously fire in cardiac muscle \cite{MartinHall1999}
and brain tissue \cite{Singer1999}.
External forcing can cause or regulate 
synchronization. % is often achieved via external forcing
%such as when
%single particle oscillators
For instance,
an electrical pacemaker %(an electrical generator)
pulses to regulate a heart beat,
and 
a flashing light can modify the
pattern of flashing fireflies .% \cite{Agrawal2013}.

%Christiaan Huygens 
%[has been adapted into many useful technologies...]
A particular form of 
synchronization is phase-locking or mode-locking,
which first appeared in the scientific literature
with 
Huygens' 1665 experiments on
the 
motions of two wall-mounted pendulum clocks.
Huygens demonstrated the tendency of
the pendula 
%two periodic oscillators in close proximity
to swing in time %, i.e. to synchronize,
due to interactions through the wall,  
no matter the original phase of the clocks \cite{Bennett2002}.
%phase locking
Dynamical systems %with multiple frequencies
exhibit phase-locking 
when oscillators with different rhythms adjust %couple
to a frequency ratio or mode 
of an integer number \cite{Bak1986}.
%Dynamical mode locking
%The phenomena
Synchronization 
is often studied with simple computational models
and can be visualized 
with phase
plots or Lissajous figures.
In phase space 
a mode-locked system is confined to 
a closed loop in a 
parametric space showing the relationship between two periodic functions
with patterns determined by the mode, % frequency ratio,
first reported in 1857 \cite{Lissajous1857}
and easily generated with a computer code 
or oscilloscope \cite{Tong1997}. 
%with a frequency generator,
%where two combined frequencies at perpendicular orientations
%({\it As we will demonstrate in phase plots in Figs. X, Y, Z})
It should be noted that oscillation phase and phase space
  refer to different uses of the word ``phase.''  Oscillation
  phase $\phi_0$ is the constant portion of the argument in a periodic function
  $\sin{(\omega t + \phi_0)}$ while phase space refers
  to plots of velocity versus position (such as $v_y$ vs. $y$)
  that are used to characterize oscillating systems.

%Given the prevalence of synchronization,
%it is the subject of many computational models.
Here we perform numerical studies 
on the synchronized dynamics
of confined particles driven over
a washboard shaped potential energy landscape.
Numerical studies %of harmonic systems
can yield insights
to the mechanics of naturally oscillating systems.
We chose this model for its
relevance to condensed matter systems
and ease of simulation.
%---------------------------------------------------------------
%kinks and nanofriction
%-----------------------------------------------------------------
%transition from synchronized to chaotic
%describe types of physical systems - particle types
%%In this work,
%we perform 
%numerical simulations of confined, driven particles
%to model a variety of physical phenomena.
%-----------------------------------------------------------------------
%paper outline
%In the following paper
In Section~\ref{sec:colloids} we describe colloid particles,
an relevent
experimental analog of our simulations.
We describe
our molecular dynamics model for a single particle in Section~\ref{sec:MD}.
%In this section, we restrict our discussion to a single particle,
%and include additional information
%for multiple particles in Section~\ref{sec:sync}.
%including various  and confining environments.
We summarize
our results 
%in Section~\ref{sec:results} 
including synchronized motion of a single confined particle
driven across a periodic landscape in 
Section~\ref{sec:one}.
In Section~\ref{sec:sync}
we 
extend the model to include multiple particles,
and show how confined 
interacting particles respond to the same applied driving force.
%including stationary propagation of high density kinks
%in Sec.~\ref{sec:kink}.
%We include code to simulate
%and visualize the landscape and dynamics in this section
%and supplementary material.
%we
%demonstrate how uniform environments and applied forces
%and create synchronized flow patterns.
%We present these results using standard tools of non-linear oscillators
%such phase diagrams of velocity versus position.
%PHASE DIAGRAMS SHOWING COUPLING, OR QUASI-PERIODIC COUPLING
%INDICATE COOPERATIVE MOTION IN THE Q1D SUBSTRATE
%In Section~\ref{sec:quasiperiod},
%we show how an aperiodic landscape modifies the particle dynamics.
%Finally we explore the transition to chaotic dynamics in 
%Section~\ref{sec:chaos}
%and conclude 
%in Section~\ref{sec:conclusion}. 
We conclude with
exercises for interested students 
in Section~\ref{sec:problems}.

@@

Complex dynamics can be studied with
colloid particles with experiments and simulation.
In experiment, colloids are typically spheres of 
plastic suspended in
highly de-ionized water or silica in suspended in organic solvent.
%experimental - colloids specifically
Because 
the spheres are a few microns in size
and move at rates of microns per second,
the position of a particle %is large and slow enough to
can be imaged as a function of time 
with a traditional optical camera.
%due to the large size and slow speed of the particles.
Individual colloids exhibit 
Brownian motion,
where particles change direction %in a seemingly random fashion
due to unseen 
collisions 
with particles comprising the suspending fluid \cite{}.
Brownian motion is a temperature and viscosity dependent
phenomena. \cite{}
[elaborate - move to associated problems?].

@@

@@

A single particle on this landscape
responds to the applied driving force
by synchronizing its motion across the landscape maxima 
to the periodic pattern of 
applied driving force.
In Fig.~\ref{fig:0} we
show the relationship between the applied force and particle
position % as a function of time.
for 
%We drive
a single particle driven across a periodic landscape. 
%along the y-direction
%({\it redundant, just refer back to eq. in Model?})
%and $V_{0y}=0.2$.
%This is illustrated in Fig.~\ref{fig:1}(a) where
%the red (blue) regions show local maxima (minima).
%
The numerical implementation of the landscape 
is calculated with Eq.~\ref{eq:ysubstrate} as 
\begin{equation}
  \label{eq:force}
  F^l_y(y) = -A_{p} \sin{(N_p \pi y / L)} 
\end{equation}
where the force is scaled with parameter $A_{p} = 0.1$ force units 
%in simulation units for force. % $E_0/a_0$.
%The substrate is described in 
%\begin{equation}
%  \label{eq:landscape_potential}
%  U_l(y) = U_{0y} \cos{(N_p \pi y / L)} 
%\end{equation}
with $N_p=20$ %troughs in the landscape
and spatial period $\lambda = 1.825 a_0$.
%
In Fig.~\ref{fig:0}(a)
we plot the applied driving force defined in Eq.~\ref{eq:drive}
as a function of time.
The time independent constants are 
$F^{AC}=0.05$ and $f=0.01$ cycles per time unit.
The temporal period of the driving force is
$T = 1/f = 100$.
To avoid transient oscillations at early times, 
the dc driving force is initially zero   
and 
slowly increased %$F^{DC}$.
at a rate of $\Delta F^{DC} = 0.01$ every $\Delta t = 4000$ integer timesteps 
to the constant value $F^{DC}=0.1$.
In Fig.~\ref{fig:0}(a) we plot $F^{DC}$
in blue to show the increase in 
$F^{DC}$ for the first $40$ time units.
In orange we show 
the combined ac and dc drive $F^D(t)$,
where 
the magnitude of the ac applied force is initially zero.
%since $\sin{(0)}=0$.
The vertical dashed lines in panels (a) and (b) coincide
with the condition $\sin{(\omega t)} = 1$,
where the applied force is maximum once $F^{DC}$ has become constant.

@@

The slope of $y/\lambda$
is proportional to the net force on the particle (See Eq.~\ref{eq:motion}).
This can be seen in 
Fig.~\ref{fig:0} %we demonstrate this relationship,
where local extrema appear in $y/\lambda$
when $F_{D}(t) \approx 0 $. 
When $F^D(t) > A_p$, a particle can 
overcome the barrier height of the landscape maxima,
and 
the particle hops between minima in the energy landscape.
Here 
the particle moves
in the positive y-direction
through $\Delta y = 2 \lambda$ in one time period $\Delta t = T$,
with 
the average velocity 
$\bar{v}_y = 2 \lambda f$ where $T = 1/f$.
%reaching a minimum position
%of $y/\lambda = 1$
%over the total period $T = 100$.
%The average velocity $\bar{v}_y$
%is the average displacement $\Delta y = y(t_0+T) - y(t_0)$
%over the period of the driving force.
%In Fig. ~\ref{fig:0}
%the average displacement during one time period is a single wavelength
%of the substrate $y(t_0+T) - y(t_0) = \lambda$.
%single DC case with figure
%dc_current 0.01
%ac_current 0.05
%ac_frequency 0.01
%fp_y 0.1
\begin{center}
\begin{figure}[h!]
\centering
\includegraphics[width=\columnwidth]{single_particle.pdf}
\caption{(a) The applied driving force $F^D(t)$ (orange) in Eq.~\ref{eq:drive}
  where the parameter $F^{DC}$ (blue) is slowly increased from zero
  to a constant value
  as described in the text.  (b) 
  The $y-$position of the driven particle
  as a function of time %of a single driven particle
  normalized by the period of the substrate $\lambda$.
  }
\label{fig:0}
\end{figure}
\end{center}

The ratio of $F^{AC}/F^{DC}$ and the barrier height $A_p$
control the rate and length of the particle hops
forward and backward in the landscape.
When $F^{AC}/F^{DC} > 1$
and $F^{DC} - F^{AC} > A_p$ the particle
moves in the negative y-direction 
during part of the cycle.
To explore the possible hopping patterns,
we sweep through a range of $F^{DC}$ for fixed $F^{AC}$ and $A_p$,
as shown in Fig.~\ref{fig:1}(b).
We keep $F^{DC}$ constant
for $\Delta t = 10^4$ timesteps
and 
sample the average velocity $\langle v_y \rangle $ 
as a function of $F^{DC}$.

We perform this sweep for several values
of $F^{AC}$ including the limiting case that the applied force
is a constant value.
When $F^{AC} = 0$, no steps appear
in the $\langle v_y \rangle $ vs $F^{DC}$ curve.

Synchronization
of the driving force with the landscape,
despite the overall magnitude
of $F^{DC}$.

@@

The particle 
achieves a variety of oscillation modes as a function of $F^{DC}$.
A mode is a periodic pattern of hops
with a constant average particle velocity, $\bar{v}_{y}$
over a range of driving forces $F^{DC}$.
We illustrate mode-locking in 
the velocity-force plot in Fig.~\ref{fig:1}(b).
Here $\langle v_{y} \rangle$ is increasing in non-uniform steps,
with a quantized height of
$\langle v_{y} \rangle = n \lambda f$,
where $n$ is an integer.  
%$\lambda = L/N_p = 36.5/20 = 1.825$ is the spatial period, or wavelength
%of the landscape,
%and $f = 0.01$ cycles per time unit.

@@

We 
show the dynamics 
in supplementary materials \cite{supp1}.

\begin{center}
\begin{figure}[h!]
\centering
\includegraphics[width=\columnwidth]{single}
\caption{(Average particle velocity in the y-direction $\langle v_{y} \rangle$ as a function of a constant driving force $F^{DC}$.}
\label{fig:1}
\end{figure}
\end{center}

%In the animation available in Ref.~\cite{supp1} the magenta dot represents the average velocity of the particle $\langle v_{y} \rangle$ at which the particle in Fig. 1(a) is moving.

%We examine
The period of the substrate
can be used to control the
intrinsic velocity of the dc driven particle,
and the applied ac drive can cause mode-locking
which appears as non-linear 
steps in the force-velocity relationship.

Fundamental models
of synchronization 
can include many coupled oscillators
or focus on 
a single particle driven across a periodic potential landscape.
Single particle studies 
are useful to understand synchronization
in the absence of collective effects.
%A single particle particle oscillator
%within a potential well is 
%much like a skateboarder in a half-pipe or
%a child on a swing
%moving back and forth. % through equilibrium.  % \cite{}.
The relationship between the driving force and
landscape can change the pattern of motion.  
With a constant or 
dc drive the landscape modulates 
the particle velocity, below some threshold  
the dc force is not strong enough to push the particle
across a potential maximum so the average velocity is zero,
a phenomena refered to as pinning \cite{Reichhardt2017}.  
%and dissipating energy from the system at all driving forces.
A sinusoidal or ac drive creates mode-locking,
where the average particle velocity
is fixed for a range of dc drive forces \cite{Reichhardt2015}.

@@

      %PRACTICAL APPLICATION
    %\bibitem{Agrawal2013} D. K. Agrawal, J. Woodhouse, and A. A. Seshia, Observation of locked phase dynamics and enhanced frequency stability in synchronized micromechanical oscillators, Physical Review Letters {\bf 111}, 084101 (2013).
